\section{\textbf{What is Portable Performance}}
\label{sec:whatispp}

The term portable performance generally means the ability to achieve a high level of performance on a variety of architectures.
%
In this case, high performance is relative to each target system~\cite{michaelwolfe2016}.
%
One important consideration, then, is the target architectures.
%

%
The top ranked machines in the world currently utilize technologies like general purpose graphics processing units (GPUs, e.g., NVIDIA Tesla in Titan), many-core co-processors (e.g., Intel Xeon Phi in Tianhe-2), and large multi-core CPUs (e.g., IBM Power, Intel Xeon in Tianhe-2 and others)~\cite{michaelwolfe2016},~\cite{top500thelist2016}. 
%
Further, future supercomputing designs may include low-power architectures (e.g., ARM), hybrid designs (e.g., AMD APU), or experimental designs (e.g., FPGA systems). 
%
Given this wide array of possible architectures, the value of portable performance has never before been higher.
%

\subsection{\textbf{Portable Performance Applications}}

In this section we will look at the meaning of portability and the uses of portable performance abstractions in different contexts.
%
Moreland et al.~\cite{moreland2015visualization} discussed the need for portable performance solutions for visualization software in their paper, ``Visualization for Exascale: Portable Performance is Critical".
%
While their context is visualization, their arguments span more than just visualization applications.
%
Currently in use scientific legacy applications today will not be adequate as they are now to run on exa-scale machines, or even the up coming peta-scale machines.
%
Applications such as these that need to be ready to run on new machines as they arrive.
%
These applications will need to look strongly at portable performance solutions in order to maintain relevance in the upcoming computing environments.

Portability does not mean the same thing to everyone.
%
In some cases, portability might be as simple as requiring minimal code changes to run on a new architecture.
%
Bosilca et al. ~\cite{bosilca2011performance} describe portability for their application as requiring no code changes to the main body of code in order to enable GPUs.
%
In this work, the only allowable changes were those in CUDA to launch kernels while the code that made up the kernels had to remain unchanged.
%
In this work they added a scheduler that could launch code regions onto different platforms based on availability and which one is the most beneficial for performance.

Du et al.~\cite{du2012cuda} describe their work as portable since they transitioned from CUDA to OpenCL.
%
The goal of their work was to understand the performance tradeoff between a more portable OpenCL implementation of an algorithm and the vendor specific GPU language CUDA.
%
In this work Du et al. showed that the performance of OpenCL was similar to CUDA for the compute intensive kernels but also had a higher overhead.
%
Also, it is important in OpenCL to account for architecture specific features or designs for optimum performance.
%

Portable performance as was laid out in Moreland et al.'s paper can be seen in recent work done in the field of ray tracing and volume rendering.
%
Larsen et al.~\cite{larsen2015ray} presented his work for a method of ray tracing consisting entirely of data parallel primitives.
%
In this work, all parallel operations are expressed using data parallel primitives, which are defined in a library.
%
The definitions of these primitives are then compiled to be CUDA, OpenMP, or serial executions.
%
The performance of this method is shown to be competitive with both of the top ray tracing libraries, OptiX from NVIDIA and Embree from Intel. 
%
In addition to ray tracing Larsen et al.~\cite{larsen2015volume} followed up this work with a volume rendering capability that uses the same principles for portable performance.
%
This work shows volume rendering performance similar to current industry standards while also maintaining portability.

Portable performance also exists outside of the visualization realm with scientific codes also utilizing the new methods being developed.
%
Rahaman et al.~\cite{rahaman2015portability} presents work on portable performance for nuclear reactor models by using the OCCA programming model.
%
OCCA, or the Open Concurrent Computing Abstraction, provides an interface into parallelism that can be compiled for different architectures.
%
Like many of the abstraction layers that will be discussed in Section~\ref{sec:abstractionLayers}, OCCA provides the ability to write code once and compile it into different known formats such as OpenMP, CUDA, or serially.
%
Rahaman et al. compared the performance of this abstraction against native implementations in order to weight the usefulness of the portability it provided.
%
Their results showed that for some cases optimal performance could be achieved simply on both a CPU and a GPU, but in other cases it took complex specialization in order to make the same kernel work well on both architectures.

Portability has become a pressing point for developing applications in today's computing environment.
%
Applications like those Rahaman studied, which preformed less than optimal on some architectures, are quite common as architecture specific optimizations become increasingly difficult to balance.
%
An added complication to this issue is the inclusion of legacy codes into the mix.
%
Since legacy codes are already developed and might require massive rewrites to take on a new architecture like a GPU, finding a portable performance solution will be critical to their future development.

\subsection{ \textbf{Abstraction Layers}}
\label{sec:abstractionLayers}

Abstraction layers are a frequent used method for achieving portable performance.
%
The key idea behind an abstraction layer is to hide the complexity of parallelism behind an abstraction.
%
Then the abstraction can handle how to parallelize a given section of code onto separate hardware architectures.
%
The remainder of this section surveys some of the known abstraction layers with a summary of their benefits and goals.

\subsection*{\textbf{OpenMP}}

Parallelism through OpenMP is achieved through the use of compiler directives, library routines, and environmental variables.
%
These are used to specify the high level parallelism for programs using the Fortran and C/C++ languages.
%
These directives, routines and variables have been expanded to include methods to describe how regions of code or data should be moved to another computing device, like an accelerator~\cite{openmp}.

Lee et. al.~\cite{lee2009openmp} describe several advantages for using OpenMP as a programming paradigm for use on a GPGPU:
\begin{itemize}
\item ``OpenMP is efficient at expressing loop-level parallelism in applications, which is an ideal target for utilizing GPU's highly parallel computing units to accelerate data-parallel computations."
\item ``The concept of a master thread and a pool of worker threads in OpenMP's fork-join model represents well the relationship between the master thread running on the host CPU and a pool of threads in a GPU device."
\item ``Incremental parallelization of applications, which is one of OpenMP's features, can add the same benefit to GPGPU programming."
\end{itemize}

By including target device directives as well as other supporting features, OpenMP is able to utilize its experiences in parallel computing and offer a familiar solution to programmers who need to make new or existing algorithms and codes work for parallel CPUs, GPUs and more~\cite{ayguade2010extending}~\cite{lee2010openmpc}.

\subsection*{\textbf{OpenACC}}

OpenACC enables the offloading of loops and regions of code onto accelerator devices.
%
The OpenACC API uses a host-directed model of execution where the main program runs on the host, or CPU, and the computational work is offloaded to a device accelerator, like a GPU.
%
The OpenACC memory model outlines two memory spaces which do not automatically synchronize, requiring explicit synchronization calls between memory spaces.
%
OpenACC operates in a similar fashion to OpenMP by using compiler directives to define regions of code for their operations to affect~\cite{wienke2012openacc}.

OpenACC is designed to be portable.
%
Its directive based programming allows programmers to create high-level host+accelerator applications without needing to explicitly handle many of the extra aspects to working on an accelerator~\cite{openacc}.

OpenACC has demonstrated the ability to achieve reasonable performance on multiple platforms.
%
Wang et. al.~\cite{wang2013performance} performed a performance study showing that for some benchmarks the OpenACC versions were able to achieve more than 82\% performance when compared with peak performance for both the Intel Knights Corner and NVIDIA Kepler architectures.

\subsection*{\textbf{Thrust}}

Thrust is a library of algorithms and data structures that can be used to provide an interface to parallel programming in order to increase a programmer's productivity.
%
Thrust is designed similar to the standard template library, allowing programmers familiar with the C++ STL to feel instantly comfortable working in the Thrust environment.
%
Through this design, Thrust lowers the barrier to entry for allowing access to GPU hardware and memory without the needed to interact with the CUDA API~\cite{hoberock2010thrust}.

In addition to adding parallel algorithms, Thrust provides multiple compilable backend technologies that allow the programmer to write their algorithms using Thrust and then compile them in CUDA, TBB, and OpenMP.
%
This enables a wide array of portable solutions that programmers can take advantage of in order to much more easily write portable and performant applications~\cite{thrust}.

Thrust offers a variety of algorithms with significant performance advantages to direct naive implementations, leading to real world performance gains.
%
Some examples of those performance gains can be seen in the implementations of the fill and radix sort algorithms.
Thrust provides a fill algorithm that produces a 32x performance gains over a naive algorithm implementations as well as a radix sort algorithm that provides a 2.7x performance gains by utilizing only significant bits when possible.
%
These performance gains come for free when using a Thrust algorithm to accomplish a data parallel task~\cite{bell2011thrust}.

In addition, Thrust provides all of the main data parallel operations defined in Guy Blelloch's work~\cite{blelloch1990vector}.
%
Blelloch's work is significant in that it provides a foundation for data parallel processing and algorithm development through a series of well defined vectorized operations.
%
One method of achieving performance is to then rewrite an algorithm using data parallel primitives or algorithms and then use the existing Thrust methods to perform the operations.

\subsection*{\textbf{RAJA}}

The RAJA portability layer is designed to be a lightweight method of providing loop-level parallelism in existing codes.
%
The idea behind the design was that, especially at institutions like LLNL, there are a large number of legacy scientific codes that will need to make some sort of transition in order to utilize upcoming architectures.
%
RAJA was designed to be able to replace current loops with a wrapper loop to at first make no change or impact. 
%
Then once the RAJA abstraction layer is in place, the loop can be changed to run on different architectures  and with different parallel modes~\cite{hornung2014raja}~\cite{hornung2016raja}.

RAJA achieved their flexibility through macro replacements in their library.
%
By changing a compile time option the user can define if they want the OpenMP parallel launcher, a CUDA kernel launcher, or a serial launcher.
%
In this manner RAJA is a useful tool for generically replacing large numbers of parallel loops with a consistent theme that creates inlined parallel code for the compilers to optimize, instead of large and sometimes convoluted template models~\cite{hornung2014raja}~\cite{hornung2016raja}.

In addition to providing a library, the RAJA project provides a second approach to portability.
%
A RAJA like approach involves simple custom macro definitions, such as making a parallel loop by replacing a for loop with a macro function.
%
Then different parallel launchers can be defined for each target architecture without changing the body of the loop, minimizing code redundancy between versions.
%
Finally, at compile time one of the architectures or versions of the parallel loop is chosen and all of the loops defined with the macro will be launched for whichever version was chosen.

\subsection*{\textbf{Kokkos}}

The Kokkos~\cite{edwards2012kokkos} C++ library provides a programming model that enables performance portability across devices.
%
The objective of the Kokkos library is to allow as much of the user's code as possible to be compiled for different devices, while obtaining the same performance as a variant of the code that was written specifically for that device.
%
Kokkos uses the idea of execution and memory spaces to provide an abstraction to the problem.
%
In their model threads are said to execute in an execution space, while data resides within a memory space.
%
Then relationships are defined between the different execution and memory spaces~\cite{edwards2014kokkos}.

Parallelism in Kokkos comes from parallel execution patterns; data parallel and task parallel patterns are used.
%
The primary data parallel patterns are: parallel\_for, parallel\_reduce, and parallel\_scan.
%
The data parallel computational kernels are implemented as standard C++ functors.
%

The Kokkos abstraction layer has demonstrated performance of approximately 90\% of the performance of the optimized architecture specific versions for kernel tests and mini-applications.
%
Kokkos has demonstrated performance on Xeon, Xeon Phi, and Kepler architectures, showing the portability of this solution~\cite{edwards2014kokkos}~\cite{edwards2012manycore}.

\subsection*{\textbf{Chapel}}

Chapel is an object-oriented parallel programming language which was designed from first principles~\cite{sidelnik2012performance}.
%
Chapel was developed in order to improve the programmability and productivity of development on parallel machines.
%
DARPA's High Performance Computing Systems defines productivity as ``a combination of performance, programmability, portability, and robustness"~\cite{chamberlain2007parallel}.
%
Chapel used this idea to make a global-view parallel language that uses a block-imperative programming style.
%
Chapel purposely avoided building on the C or Fortran languages in order to help programmers avoid falling back into sequential programming patterns~\cite{chamberlain2007parallel}.

Chapel uses a code generation design to generate parallel C or CUDA code.
%
The Chapel language defines the parallelism and so can be used as the basis for optimized code generation on many different platforms.
%
Chapel uses this design to achieve portability and performance with their language.

Chapel's design goal is to support any parallel algorithm that a programmer could conceive without the need to fall back to other parallel libraries.
%
Chapel supports concepts for describing parallelism separate from those used to describe locality.
%
It supports programming at higher and lower levels, as well as providing advanced higher-level features such as data distributions or parallel loop schedules~\cite{bradchamberlain2013}.


\subsection*{\textbf{VTK-m}}

VTK-m is the result of a collaboration between three separate groups and three separate national labs coming together and joining forces with Kitware, the primary maintainers of the current VTK (Visualization ToolKit) software.
%
Visualization applications use VTK in order to express visualization algorithms and data structures in their codes.
%
VTK-m came about from the three projects, EAVL, DAX, and PISTON, with the design goal of being a portable performance solution for visualization applications and algorithms.
%

The VTK-m framework takes the concepts of data parallel primitives and patterns generated from those primitives to provide a framework for accomplishing visualization algorithms.
%
These data parallel primitives can be compiled for different platforms, helping VTK-m achieve portable performance~\cite{Moreland:CGA2016}~\cite{JSFI77}.

The contributions of the three projects to VTK-m are as follows:
\begin{itemize}
\item EAVL -- Provided a robust data model. 
\item DAX -- Provided a model for parallel work dispatching. 
\item Piston -- Provided many data parallel algorithms and implementations.
\end{itemize}


\subsubsection*{\textbf{EAVL}}

EAVL, or the Extreme-scale Analysis and Visualization Library~\cite{jeremyEAVL}, was developed with three goals in mind:
\begin{itemize}
\item A flexible data model -- ``Expanding on traditional models to support current and forthcoming scientific data sets."
\item High parallel efficiency -- ``Improve memory and algorithmic efficiency through the enhanced data model, and support stricter memory controls and accelerator device memory models."
\item Scalability -- ``Support distributed and data parallelism, and transparently target heterogeneous systems."
\end{itemize}
%

\subsubsection*{\textbf{Dax}}

Dax, or Data Analysis at Extreme, is a library developed to support fine grained concurrency for data analysis and visualization algorithms.
%
This library provides a dispatcher that schedules worklets onto data items.
%
The Dax toolkit simplifies the development of parallel visualization algorithms and provides a data parallel framework for scheduling and launching parallel jobs~\cite{morelanddax}~\cite{moreland2011dax}.

\subsubsection*{\textbf{PISTON}} 

The Portable Data-Parallel Visualization and Analysis Library~\cite{lo2012piston}, referred to as PISTON, is a cross-platform library that provides operations for scientific visualization and analysis.
%
These operations are performed using data parallel primitives and the NVIDIA Thrust library.
%
PISTON uses Thrust to perform the data parallel operations and for its cross-platform compatibility.
%
PISTON adds useful algorithms for data visualization and analysis as well as an interface into the Thrust calls~\cite{PISTON}.
